Physical Chemistry Assignment Four

### 2015-11-28 classical/latex/a8.tex

\section{Problem One: }
\textbf{A pendulum consists of a mass m suspended by a massless spring with un-extended length $b$ and spring constant $k$. Find Hamilton’s equations of motion.}

\section{Problem Two: }
\textbf{The potential for an non-harmonic oscillator is:}

@@

\section{Problem Three: }
\textbf{Reconsider the problem of the two masses and three springs (seen in class) in the event that the three springs all have different force constants (but the masses are equal). Find the two characteristic frequencies and compare the magnitudes with the natural frequencies of the two oscillators in the absence of coupling.}

@@

\end{document}

>>>>

\section{Problem One: Springy Pendulum}
\textbf{A pendulum consists of a mass m suspended by a massless spring with un-extended length $b$ and spring constant $k$. Find Hamilton’s equations of motion.}

\begin{align*}
  x &= (b+\ell)\sin\theta\\
  y &= -(b+\ell)\cos\theta\\
  \dot{x} &= \dot{\ell}\sin\theta + (b+\ell)\cos\theta\dot{\theta}\\
  \dot{y} &= -\dot{\ell}\cos\theta + (b+\ell)\sin\theta\dot{\theta}\\
  \vspace{1cm}\\
  T &= \frac{1}{2}m\left(\dot{x}^2 + \dot{y}^2\right)\\
  &= \frac{1}{2}m\left(\dot{\ell}^2 + (b+\ell)^2\dot{\theta}^2\right)\\
  U &= mgy = -mg(b + \ell)\cos\theta + \frac{1}{2}k\ell^2\\
  \mathcal{L} &= \frac{1}{2}m\left(\dot{\ell}^2 + (b+\ell)^2\dot{\theta}^2\right) + mg(b + \ell)\cos\theta - \frac{1}{2}k\ell^2\\
  \vspace{1cm}\\
  p_{\ell} &= \frac{d\mathcal{L}}{d\dot{\ell}} = m\dot{\ell}\\
  p_{\theta} &= \frac{d\mathcal{L}}{d\dot{\theta}} = m(b + \ell)\dot{\theta}\\
  \mathcal{H} &= m\dot{\ell}^2 + m(b + \ell)\dot{\theta}^2 - \frac{1}{2}m\left(\dot{\ell}^2 + (b+\ell)^2\dot{\theta}^2\right) - mg(b + \ell)\cos\theta + \frac{1}{2}k\ell^2\\
  &= \frac{p_{\ell}^2}{m} + \frac{p_{\theta}^2}{m(b + \ell)} - \frac{1}{2}m\left(\frac{p_{\ell}^2}{m^2} + \frac{p_{\theta}^2}{m^2}\right) - mg(b + \ell)\cos\theta + \frac{1}{2}k\ell^2\\
  \vspace{1cm}\\
  \dot{p_{\theta}} &= -\frac{d\mathcal{H}}{d\theta}\\
  &= m\ddot{\theta} = -mg(b + \ell)\sin\theta\\
  \dot{p_{\ell}} &= -\frac{d\mathcal{H}}{d\ell}\\
  &= m\ddot{\ell} = -m\dot{\theta}^2 + m\left(b+\ell\right)\dot{\theta}^2 + mg\cos\theta + 2k\ell\\
\end{align*}

This was an interesting usage of Hamiltonian mechanics to solve a very weird physical system. This would have been a difficult problem to solve using Newtonian mechanics. It probably would have been easier to solve with Lagrangian mechanics due to there being fewer steps, but Hamiltonian worked well. The end result is a set of coupled motion equations with no obvious solution, which is in line with what we were told about in class.

\section{Problem Two: Non-harmonic Oscillator}
\textbf{The potential for an non-harmonic oscillator is:}

@@

\begin{align*}
  T &= \frac{1}{2}m\dot{x}^2\\
  \mathcal{L} &= \frac{1}{2}m\dot{x}^2 - \frac{kx^2}{2} - \frac{bx^4}{4}\\
  \vspace{1cm}
  p &= \frac{d\mathcal{L}}{d\dot{x}} = m\dot{x}\\
  \mathcal{H} &= m\dot{x} - \frac{1}{2}m\dot{x}^2 + \frac{kx^2}{2} + \frac{bx^4}{4}\\
  \dot{p} &= -\frac{d\mathcal{H}}{dx} = -kx - bx^3\\
\end{align*}

We combine this equation with our knowledge of the definition of momentum $p = m\dot{x}$ to find the Hamilton Equation of Motion:

\begin{align*}
  m\ddot{x} &= -kx - bx^3\\
\end{align*}

\section{Problem Three: Two masses Three springs}
\textbf{Reconsider the problem of the two masses and three springs (seen in class) in the event that the three springs all have different force constants (but the masses are equal). Find the two characteristic frequencies and compare the magnitudes with the natural frequencies of the two oscillators in the absence of coupling.}

\begin{align*}
  T &= \frac{1}{2}m\dot{x}^2 + \frac{1}{2}m\dot{y}^2\\
  U &= \frac{1}{2}k_1x^2 + \frac{1}{2}k_3y^2 + \frac{1}{2}k_2\left(y - x\right)^2\\
  \mathcal{L} &= \frac{1}{2}\left(k_1x^2 + k_3y^2 + k_2\left(y-x\right)^2 + m\dot{x}^2 + m\dot{y}^2\right)\\
\end{align*}

By solving the Euler-Lagrange equation for this Lagrangian we find the following motion equations:

\begin{align*}
  m\ddot{x} &= k_1x - k_2\left(y-x\right)\\
  m\ddot{y} &= k_3y + k_2\left(y-x\right)\\
\end{align*}

We transfer these motion equations to matrix form and solve their determinant:

\begin{align*}
  det
  \begin{pmatrix}
    k_1+k_2-\omega^2m & -k_2\\
    -k_2 & k_2 + k_3-\omega^2m\\
  \end{pmatrix}
  = 0\\
  \left(k_1+k_2-\omega^2m\right)\left(k_2 + k_3 - \omega^2m\right) - k_2^2 = 0\\
  &\omega^4m^2 + \omega^2m\left(-k_1-2k_2-k_3\right) + \left(k_1k_2 + k_1k_3 + k_2k_3\right) = 0\\
  \omega^2 &= \frac{m\left(k_1+2k_2+k_3\right) \pm \sqrt{m^2\left(k_1+2k_2+k_3\right)^2 - 4m^2\left(k_1k_2+k_1k_3+k_2k_3\right)}}{2m^2}\\
  \omega_1 &= \sqrt{\frac{m\left(k_1+2k_2+k_3\right) + \sqrt{m^2\left(k_1+2k_2+k_3\right)^2 - 4m^2\left(k_1k_2+k_1k_3+k_2k_3\right)}}{2m^2}}\\
  &= \sqrt{\frac{\left(k_1+2k_2+k_3\right) + \sqrt{k_1^2+4k_2^2+k_3^2-2k_1k_3}}{2m}}\\
  \omega_2 &= \sqrt{\frac{m\left(k_1+2k_2+k_3\right) - \sqrt{m^2\left(k_1+2k_2+k_3\right)^2 - 4m^2\left(k_1k_2+k_1k_3+k_2k_3\right)}}{2m^2}}\\
  &= \sqrt{\frac{k_1+2k_2+k_3 - \sqrt{k_1^2+4k_2^2+k_3^2-2k_1k_3}}{2m}}\\
\end{align*}

We can compare the uncoupled frequencies (where $k_2=0$) to these:

\begin{align*}
  \lim_{k_2\rightarrow0}\omega_1 &= \sqrt{\frac{\left(k_1+k_3\right) + \sqrt{k_1^2+k_3^2-2k_1k_3}}{2m}} = \sqrt{\frac{k_1}{m}}\\
  \lim_{k_2\rightarrow0}\omega_2 &= \sqrt{\frac{\left(k_1+k_3\right) - \sqrt{k_1^2+k_3^2-2k_1k_3}}{2m}} = \sqrt{\frac{k_3}{m}}\\
\end{align*}

This is just what we expect, a system of two independent uncoupled harmonic oscillators.\\

This was a neat problem. There wasn't really anything different about it compared with other similar problems we've done, except perhaps the determinant equation being too difficult to factor, so having to use the quadratic formula on a fourth-order polynomial. It was nice that the uncoupled limit worked out so well.\\

@@

We begin by defining our generalized coordinates. $\Theta$ is the hoop's CM-connection point angle with the horizontal and $\phi$ is the mass-hoop CM angle with respect to the horizontal.

\begin{align*}
  x &= R\sin\theta + R\sin\phi\\
  y &= -R\cos\theta + -R\cos\phi\\
  \dot{x}^2 + \dot{y}^2 &= R^2\left(\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\theta}\dot{\phi}\cos\left(\theta-\phi\right)\right)\\
  \vspace{1cm}\\
  T &= \frac{1}{2}M\left(\dot{x}^2+\dot{y}^2\right) + \frac{1}{2}I\omega^2\\
  &= \frac{1}{2}MR^2\left(\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\phi}\dot{\theta}\cos(\theta-\phi) + 2\dot{\theta}^2\right)\\
  \mathcal{L} &= \frac{1}{2}MR^2\left(\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\phi}\dot{\theta}\cos(\theta-\phi) + 2\dot{\theta}^2\right) + mgR\left(\cos\theta + \cos\phi\right) + mgR\cos\theta\\
  &= \frac{1}{2}MR^2\left(3\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\phi}\dot{\theta}\cos(\theta-\phi)\right) + mgR\left(2\cos\theta + \cos\phi\right)\\
  &\hspace{1cm} cos(\theta) = 1 - \frac{\theta^2}{2},\hspace{1cm} cos(\phi) = 1 - \frac{\phi^2}{2}\\
  &= \frac{1}{2}MR^2\left(3\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\phi}\dot{\theta}\right) + mgR\left(3 - \theta^2 - \frac{\phi^2}{2}\right)\\
  \vspace{1cm}\\
  \frac{d}{dt}\frac{d\mathcal{L}}{d\dot{\theta}} &= \frac{d\mathcal{L}}{d\theta}\\
  \frac{1}{2}MR^2\left(6\ddot{\theta} + 2\ddot{\phi}\right) &= -2MgR\theta\\
  R\left(6\ddot{\theta} + 2\ddot{\phi}\right) &= -2g\theta\\
  3\ddot{\theta} + \ddot{\phi} &= \frac{-2g}{R}\theta\\
  \frac{d}{dt}\frac{d\mathcal{L}}{d\dot{\phi}} &= \frac{d\mathcal{L}}{d\phi}\\
  \frac{1}{2}MR^2\left(2\ddot{\phi} + 2\ddot{\theta}\right) &= -MgR\phi\\
  \ddot{\phi} + \ddot{\theta} &= \frac{-g}{R}\phi\\
\end{align*}

We have obtained our equations of motion using Lagrangian mechanics. We now guess the solution using the complex exponential function. This will give us a system of two equations and three unknowns. However, this particular system allows us to cancel out the A and B coefficients and solve for $\omega^2$.\\

\begin{align*}
  \theta &= Ae^{i\omega t}\\
  \phi &= Be^{i\omega t}\\
  3\omega^2A + \omega^2B &= \frac{-2g}{R}A\\
  \omega^2A + \omega^2B &= \frac{-g}{R}B\\
  A = B\left(\frac{g}{R\omega^2}-1\right)\\
  3\omega^2\left(\frac{g}{R\omega^2}-1\right) + \omega^2 &= \frac{2g}{R}\left(\frac{g}{R\omega^2}-1\right)\\
  \frac{3g}{R} - 3\omega^2 + \omega^2 = \frac{2g^2}{R^2}\frac{1}{\omega^2} - \frac{2g}{R}\\
  -2\omega^4 + \frac{5g}{R}\omega^2 - \frac{2g^2}{R^2} &= 0\\
  \omega^2&= \frac{-\frac{5g}{R} \pm \sqrt{\frac{25g^2}{R^2} - 8\frac{2g^2}{R^2}}}{-4} = \frac{g}{2R}, \frac{2g}{R}\\
  \omega_1 &= \sqrt{\frac{g}{2R}}\\
  \omega_2 &= \sqrt{\frac{2g}{R}}\\
\end{align*}

This was a very cool problem. We applied a bunch of different physics to solve it: Taylor approximations to the small-angle cosine, Lagrangian mechanics, and a complex set of generalized coordinates. This was a long and difficult problem, but I enjoyed thinking about how to set it up. The trick of using the small-angle approximation as a Taylor series, but only applying it to the large terms (not the terms involving the angular velocities) was very tricky. I think that thinking about how values in your equations influence the behavior of the equation, instead of just operating on them, is important to doing good physics.\\
\end{document}

### 2015-11-28 classical/latex/a9.tex

\title{Classical Mechanics Assignment 8}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

@@

\section{Problem One: Springy Pendulum}
\textbf{A pendulum consists of a mass m suspended by a massless spring with un-extended length $b$ and spring constant $k$. Find Hamilton’s equations of motion.}

\begin{align*}
  x &= (b+\ell)\sin\theta\\
  y &= -(b+\ell)\cos\theta\\
  \dot{x} &= \dot{\ell}\sin\theta + (b+\ell)\cos\theta\dot{\theta}\\
  \dot{y} &= -\dot{\ell}\cos\theta + (b+\ell)\sin\theta\dot{\theta}\\
  \vspace{1cm}\\
  T &= \frac{1}{2}m\left(\dot{x}^2 + \dot{y}^2\right)\\
  &= \frac{1}{2}m\left(\dot{\ell}^2 + (b+\ell)^2\dot{\theta}^2\right)\\
  U &= mgy = -mg(b + \ell)\cos\theta + \frac{1}{2}k\ell^2\\
  \mathcal{L} &= \frac{1}{2}m\left(\dot{\ell}^2 + (b+\ell)^2\dot{\theta}^2\right) + mg(b + \ell)\cos\theta - \frac{1}{2}k\ell^2\\
  \vspace{1cm}\\
  p_{\ell} &= \frac{d\mathcal{L}}{d\dot{\ell}} = m\dot{\ell}\\
  p_{\theta} &= \frac{d\mathcal{L}}{d\dot{\theta}} = m(b + \ell)\dot{\theta}\\
  \mathcal{H} &= m\dot{\ell}^2 + m(b + \ell)\dot{\theta}^2 - \frac{1}{2}m\left(\dot{\ell}^2 + (b+\ell)^2\dot{\theta}^2\right) - mg(b + \ell)\cos\theta + \frac{1}{2}k\ell^2\\
  &= \frac{p_{\ell}^2}{m} + \frac{p_{\theta}^2}{m(b + \ell)} - \frac{1}{2}m\left(\frac{p_{\ell}^2}{m^2} + \frac{p_{\theta}^2}{m^2}\right) - mg(b + \ell)\cos\theta + \frac{1}{2}k\ell^2\\
  \vspace{1cm}\\
  \dot{p_{\theta}} &= -\frac{d\mathcal{H}}{d\theta}\\
  &= m\ddot{\theta} = -mg(b + \ell)\sin\theta\\
  \dot{p_{\ell}} &= -\frac{d\mathcal{H}}{d\ell}\\
  &= m\ddot{\ell} = -m\dot{\theta}^2 + m\left(b+\ell\right)\dot{\theta}^2 + mg\cos\theta + 2k\ell\\
\end{align*}

This was an interesting usage of Hamiltonian mechanics to solve a very weird physical system. This would have been a difficult problem to solve using Newtonian mechanics. It probably would have been easier to solve with Lagrangian mechanics due to there being fewer steps, but Hamiltonian worked well. The end result is a set of coupled motion equations with no obvious solution, which is in line with what we were told about in class.

\section{Problem Two: Non-harmonic Oscillator}
\textbf{The potential for an non-harmonic oscillator is:}

\begin{equation*}
  U(x) = \frac{kx^2}{2} + \frac{bx^4}{4}\\
\end{equation*}

\textbf{where k and b are positive constants. Find Hamilton’s equations of motion. No final discussion needed.}

\begin{align*}
  T &= \frac{1}{2}m\dot{x}^2\\
  \mathcal{L} &= \frac{1}{2}m\dot{x}^2 - \frac{kx^2}{2} - \frac{bx^4}{4}\\
  \vspace{1cm}
  p &= \frac{d\mathcal{L}}{d\dot{x}} = m\dot{x}\\
  \mathcal{H} &= m\dot{x} - \frac{1}{2}m\dot{x}^2 + \frac{kx^2}{2} + \frac{bx^4}{4}\\
  \dot{p} &= -\frac{d\mathcal{H}}{dx} = -kx - bx^3\\
\end{align*}

We combine this equation with our knowledge of the definition of momentum $p = m\dot{x}$ to find the Hamilton Equation of Motion:

\begin{align*}
  m\ddot{x} &= -kx - bx^3\\
\end{align*}

\section{Problem Three: Two masses Three springs}
\textbf{Reconsider the problem of the two masses and three springs (seen in class) in the event that the three springs all have different force constants (but the masses are equal). Find the two characteristic frequencies and compare the magnitudes with the natural frequencies of the two oscillators in the absence of coupling.}

\begin{align*}
  T &= \frac{1}{2}m\dot{x}^2 + \frac{1}{2}m\dot{y}^2\\
  U &= \frac{1}{2}k_1x^2 + \frac{1}{2}k_3y^2 + \frac{1}{2}k_2\left(y - x\right)^2\\
  \mathcal{L} &= \frac{1}{2}\left(k_1x^2 + k_3y^2 + k_2\left(y-x\right)^2 + m\dot{x}^2 + m\dot{y}^2\right)\\
\end{align*}

By solving the Euler-Lagrange equation for this Lagrangian we find the following motion equations:

\begin{align*}
  m\ddot{x} &= k_1x - k_2\left(y-x\right)\\
  m\ddot{y} &= k_3y + k_2\left(y-x\right)\\
\end{align*}

We transfer these motion equations to matrix form and solve their determinant:

\begin{align*}
  det
  \begin{pmatrix}
    k_1+k_2-\omega^2m & -k_2\\
    -k_2 & k_2 + k_3-\omega^2m\\
  \end{pmatrix}
  = 0\\
  \left(k_1+k_2-\omega^2m\right)\left(k_2 + k_3 - \omega^2m\right) - k_2^2 = 0\\
  &\omega^4m^2 + \omega^2m\left(-k_1-2k_2-k_3\right) + \left(k_1k_2 + k_1k_3 + k_2k_3\right) = 0\\
  \omega^2 &= \frac{m\left(k_1+2k_2+k_3\right) \pm \sqrt{m^2\left(k_1+2k_2+k_3\right)^2 - 4m^2\left(k_1k_2+k_1k_3+k_2k_3\right)}}{2m^2}\\
  \omega_1 &= \sqrt{\frac{m\left(k_1+2k_2+k_3\right) + \sqrt{m^2\left(k_1+2k_2+k_3\right)^2 - 4m^2\left(k_1k_2+k_1k_3+k_2k_3\right)}}{2m^2}}\\
  &= \sqrt{\frac{\left(k_1+2k_2+k_3\right) + \sqrt{k_1^2+4k_2^2+k_3^2-2k_1k_3}}{2m}}\\
  \omega_2 &= \sqrt{\frac{m\left(k_1+2k_2+k_3\right) - \sqrt{m^2\left(k_1+2k_2+k_3\right)^2 - 4m^2\left(k_1k_2+k_1k_3+k_2k_3\right)}}{2m^2}}\\
  &= \sqrt{\frac{k_1+2k_2+k_3 - \sqrt{k_1^2+4k_2^2+k_3^2-2k_1k_3}}{2m}}\\
\end{align*}

We can compare the uncoupled frequencies (where $k_2=0$) to these:

\begin{align*}
  \lim_{k_2\rightarrow0}\omega_1 &= \sqrt{\frac{\left(k_1+k_3\right) + \sqrt{k_1^2+k_3^2-2k_1k_3}}{2m}} = \sqrt{\frac{k_1}{m}}\\
  \lim_{k_2\rightarrow0}\omega_2 &= \sqrt{\frac{\left(k_1+k_3\right) - \sqrt{k_1^2+k_3^2-2k_1k_3}}{2m}} = \sqrt{\frac{k_3}{m}}\\
\end{align*}

This is just what we expect, a system of two independent uncoupled harmonic oscillators.\\

This was a neat problem. There wasn't really anything different about it compared with other similar problems we've done, except perhaps the determinant equation being too difficult to factor, so having to use the quadratic formula on a fourth-order polynomial. It was nice that the uncoupled limit worked out so well.\\

\section{Problem Four: }
\textbf{A thin hoop of radius R and mass M oscillates in its own plane hanging from a single fixed point. Attached to the hoop is a small mass M constrained to move (in a frictionless manner) along the hoop. The inertia of the loop for rotation around a point on the loop is $I=2MR^2$. Consider only small oscillations, and show that the eigenfrequencies are:}

\begin{equation*}
  \omega_1 = \sqrt{\frac{2g}{R}}
\end{equation*}

\begin{equation*}
  \omega_2 = \sqrt{\frac{g}{2R}}
\end{equation*}

We begin by defining our generalized coordinates. $\Theta$ is the hoop's CM-connection point angle with the horizontal and $\phi$ is the mass-hoop CM angle with respect to the horizontal.

\begin{align*}
  x &= R\sin\theta + R\sin\phi\\
  y &= -R\cos\theta + -R\cos\phi\\
  \dot{x}^2 + \dot{y}^2 &= R^2\left(\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\theta}\dot{\phi}\cos\left(\theta-\phi\right)\right)\\
  \vspace{1cm}\\
  T &= \frac{1}{2}M\left(\dot{x}^2+\dot{y}^2\right) + \frac{1}{2}I\omega^2\\
  &= \frac{1}{2}MR^2\left(\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\phi}\dot{\theta}\cos(\theta-\phi) + 2\dot{\theta}^2\right)\\
  \mathcal{L} &= \frac{1}{2}MR^2\left(\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\phi}\dot{\theta}\cos(\theta-\phi) + 2\dot{\theta}^2\right) + mgR\left(\cos\theta + \cos\phi\right) + mgR\cos\theta\\
  &= \frac{1}{2}MR^2\left(3\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\phi}\dot{\theta}\cos(\theta-\phi)\right) + mgR\left(2\cos\theta + \cos\phi\right)\\
  &\hspace{1cm} cos(\theta) = 1 - \frac{\theta^2}{2},\hspace{1cm} cos(\phi) = 1 - \frac{\phi^2}{2}\\
  &= \frac{1}{2}MR^2\left(3\dot{\theta}^2 + \dot{\phi}^2 + 2\dot{\phi}\dot{\theta}\right) + mgR\left(3 - \theta^2 - \frac{\phi^2}{2}\right)\\
  \vspace{1cm}\\
  \frac{d}{dt}\frac{d\mathcal{L}}{d\dot{\theta}} &= \frac{d\mathcal{L}}{d\theta}\\
  \frac{1}{2}MR^2\left(6\ddot{\theta} + 2\ddot{\phi}\right) &= -2MgR\theta\\
  R\left(6\ddot{\theta} + 2\ddot{\phi}\right) &= -2g\theta\\
  3\ddot{\theta} + \ddot{\phi} &= \frac{-2g}{R}\theta\\
  \frac{d}{dt}\frac{d\mathcal{L}}{d\dot{\phi}} &= \frac{d\mathcal{L}}{d\phi}\\
  \frac{1}{2}MR^2\left(2\ddot{\phi} + 2\ddot{\theta}\right) &= -MgR\phi\\
  \ddot{\phi} + \ddot{\theta} &= \frac{-g}{R}\phi\\
\end{align*}

We have obtained our equations of motion using Lagrangian mechanics. We now guess the solution using the complex exponential function. This will give us a system of two equations and three unknowns. However, this particular system allows us to cancel out the A and B coefficients and solve for $\omega^2$.\\

\begin{align*}
  \theta &= Ae^{i\omega t}\\
  \phi &= Be^{i\omega t}\\
  3\omega^2A + \omega^2B &= \frac{-2g}{R}A\\
  \omega^2A + \omega^2B &= \frac{-g}{R}B\\
  A = B\left(\frac{g}{R\omega^2}-1\right)\\
  3\omega^2\left(\frac{g}{R\omega^2}-1\right) + \omega^2 &= \frac{2g}{R}\left(\frac{g}{R\omega^2}-1\right)\\
  \frac{3g}{R} - 3\omega^2 + \omega^2 = \frac{2g^2}{R^2}\frac{1}{\omega^2} - \frac{2g}{R}\\
  -2\omega^4 + \frac{5g}{R}\omega^2 - \frac{2g^2}{R^2} &= 0\\
  \omega^2&= \frac{-\frac{5g}{R} \pm \sqrt{\frac{25g^2}{R^2} - 8\frac{2g^2}{R^2}}}{-4} = \frac{g}{2R}, \frac{2g}{R}\\
  \omega_1 &= \sqrt{\frac{g}{2R}}\\
  \omega_2 &= \sqrt{\frac{2g}{R}}\\
\end{align*}

This was a very cool problem. We applied a bunch of different physics to solve it: Taylor approximations to the small-angle cosine, Lagrangian mechanics, and a complex set of generalized coordinates. This was a long and difficult problem, but I enjoyed thinking about how to set it up. The trick of using the small-angle approximation as a Taylor series, but only applying it to the large terms (not the terms involving the angular velocities) was very tricky. I think that thinking about how values in your equations influence the behavior of the equation, instead of just operating on them, is important to doing good physics.\\
\end{document}

>>>>

\title{Classical Mechanics Assignment 9}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

@@

\section{Problem One: }

\end{document}

### 2015-11-28 physical_chemistry/latex/a4.tex
add
Add stuffs.

b.)
\begin{align*}
  \Delta H &= -T^2\frac{\partial (\Delta G/T)}{\partial T}\\
  -\frac{\Delta H}{T^2} &= \frac{\partial (\Delta G/T)}{\partial T}\\
  \int_{T_1}^{T_2} &\left(-\frac{\Delta H_1 + (T-T_1)\Delta C_P}{T^2}\right)dT = \frac{\Delta G}{T}\Bigg|_{\frac{\Delta G_1}{T_1}}^{\frac{\Delta G_2}{T_2}}\\
  &\Delta H_1\left(\frac{1}{T_2}-\frac{1}{T_1}\right) + \ln\left(\frac{T_1}{T_2}\right)\Delta C_P - T_1\Delta C_P\left(\frac{1}{T_2}-\frac{1}{T_1}\right) = \frac{\Delta G_2}{T_2} - \frac{\Delta G_1}{T_1}\\
  \Delta G_2 &= \left(\Delta H_1 - T_1\Delta C_P\right)\left(1 - \frac{T_2}{T_1}\right) + T_2\ln\left(\frac{T_1}{T_2}\right)\Delta C_P + \frac{\Delta G_1}{T_1}\\
\end{align*}

@@

>>>>

b.)
\begin{align*}
  \Delta H &= -T^2\frac{\partial (\Delta G/T)}{\partial T}\\
  -\frac{\Delta H}{T^2} &= \frac{\partial (\Delta G/T)}{\partial T}\\
  \int_{T_1}^{T_2} &\left(-\frac{\Delta H_1 + (T-T_1)\Delta C_P}{T^2}\right)dT = \frac{\Delta G}{T}\Bigg|_{\frac{\Delta G_1}{T_1}}^{\frac{\Delta G_2}{T_2}}\\
  &\Delta H_1\left(\frac{1}{T_2}-\frac{1}{T_1}\right) + \ln\left(\frac{T_1}{T_2}\right)\Delta C_P - T_1\Delta C_P\left(\frac{1}{T_2}-\frac{1}{T_1}\right) = \frac{\Delta G_2}{T_2} - \frac{\Delta G_1}{T_1}\\
  \Delta G_2 &= \left(\Delta H_1 - T_1\Delta C_P\right)\left(1 - \frac{T_2}{T_1}\right) + T_2\ln\left(\frac{T_1}{T_2}\right)\Delta C_P + \frac{T_2\Delta G_1}{T_1}\\
\end{align*}

@@

\section{4.16}
\begin{align*}
  G &= aT + b + \frac{c}{T}\\
  S &= -\left(\frac{\partial G}{\partial T}\right)_P = \frac{c}{T^2} - a\\
  H &= G + TS = b + \frac{2c}{T}\\
\end{align*}

\section{4.17}
First we take the water from atmospheric pressure to its vapor pressure, so we can do the gas transition reversibly. This transition is $\Delta G_1$. We then convert it to a gas, this is $\Delta G_2$. We then bring the gas to the vapor pressure of the solid, this is $\Delta G_3$. We then sublimate the gas, this is $\Delta G_4$. Finally for $\Delta G_5$ we take the solid back up to the starting pressure.
\begin{align*}
  V_l &= 1.8*10^{-5} m^3\\
  V_s &= 1.96*10^{-5} m^3\\
  \Delta G_1 &= \int VdP = 1.8*10^{-5} m^3 * \left(489Pa - 10^5 Pa\right) = -1.79J\\
  \Delta G_2 &= 0 \hspace{1cm}\mbox{State transitions @ vapor pressure isenthalpic}\\
  \Delta G_3 &= \int VdP = \int \frac{RT}{P}dP = (8.314)(298)\ln\left(\frac{475}{489}\right) = -65.2J\\
  \Delta G_4 &= 0 \hspace{1cm}\mbox{State transitions @ vapor pressure isenthalpic}\\
  \Delta G_5 &= \int VdP = 1.96*10^{-5} m^3 * \left(10^5Pa - 475Pa\right) = 1.95J\\
  \Delta G &= -1.79J + -65.2J + 1.95J = -65.04\frac{J}{mol}\\
\end{align*}

\section{4.24}
a.)
\begin{align*}
  f &= Pe^{\int_0^p \frac{Z-1}{p}dp}\\
  &= Pe^{\int_0^p \frac{B}{RT}dp}\\
  &= Pe^{\frac{Bp}{RT}}\\
  &= Pe^{Z-1}\\
\end{align*}

b.)
\begin{align*}
  f_{H_2} &= PZ = P + \frac{1}{RT}\left(b - \frac{a}{RT}\right)P^2\\
  &= 50 \mbox{bar} + \frac{1}{\left(8.314*10^{-5} \frac{m^3 \mbox{bar}}{\mbox{K mol}}\right)298K}\left(0.026  - \frac{0.247}{\left(8.314*10^{-5} \frac{m^3 \mbox{bar}}{\mbox{K mol}}\right)298K}\right)\left(2500 \mbox{bar}^2\right)\\
  &= 52 bar\\
\end{align*}

\section{4.27}
To separate, take each gas and separate it. This is a zero-energy process. Then take each separated gas back to 1 mol. Use the ideal gas approximation:

\begin{align*}
  w &= -2\int pdV = -2nRT \int \frac{1}{V}dV = -2nRT\ln\left(\frac{V_2}{V_1}\right)\\
  &= 2(8.314)(298)\ln(2) = 3.4kJ\\
\end{align*}

\section{4.50}
\begin{align*}
  Z &= 1 + \left(b - \frac{a}{RT}\right)\frac{P}{RT}\\
  V &= \frac{RT}{P} + b - \frac{a}{RT}\\
  \Delta G &= \int_{P_1}^{P_2} dG = \int_{P_1}^{P_2} \frac{\partial G}{\partial P} dP\\
  &= \int_{P_1}^{P_2} VdP - \int_{P_1}^{P_2} \left(\frac{RT}{P} + b - \frac{a}{RT}\right) dP\\
  &= RT\ln\left(\frac{P_2}{P_1}\right) + \left(b-\frac{a}{RT}\right)\left(P_2-P_1\right)\\
\end{align*}

\section{4.54}
\begin{align*}
  \Delta S &= -R\sum n_i\ln y_i\\
  &= -\left(8.314\right)\left(\ln\left(\frac{1}{3}\right) + 2\ln\left(\frac{2}{3}\right)\right) = 15.87 \frac{J}{K}\\
  \Delta G &= -TS = -298*18.87 = -4.7kJ\\
\end{align*}
